Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem

### 2015-11-23 acta_manuscript.tex
added more detail on geo transfer

%% Group authors per affiliation:
\author{Shankar Kulumani\fnref{kulumanifootnote}}
\ead{skulumani@gwu.edu}
\address{George Washington University, 800 22nd St NW, Washington, DC 20052}

\author{Taeyoung Lee\fnref{leefootnote}}
\ead{tylee@gwu.edu}
\address{George Washington University, 800 22nd St NW, Washington, DC 20052}

@@

\begin{abstract}
A computational approach is developed for the design of continuous low thrust transfers in the planar circular restricted three-body problem.
The transfer design method of invariant manifolds is extended with the addition of continuous low thrust propulsion.
A reachable region is generated and it is used to determine transfer opportunities, analogous to the intersection of invariant manifolds.
The reachable set is developed on a lower dimensional \Poincare section and used to design transfer trajectories. 
%Computational geometric mechanics allows for the development of variational integrators and a discrete optimal control problem.
This is solved numerically as a discrete optimal control problem using a variational integrator.
This provides for a geometrically exact and numerically efficient method for the motion in the three-body problem.
A numerical simulation is provided developing a transfer from a \( L_1 \) periodic orbit in  the Earth-Moon system to a target orbit about the Moon.
\end{abstract}

@@

A discrete optimal control problem is formulated to determine the reachability set on a \Poincare section.
Given an initial condition and fixed time horizon, the reachable set is the set of states attainable, subject to the operational constraints of the spacecraft. 
Generation of this reachability set allows for the extension of the previous control-free missions in the three-body problem.
In addition, the generation of the reachable set allows for a more systematic method of determining initial conditions and eases the burden on the designer. 
The addition of low-thrust propulsion affords the ability to enlarge the reachable set as compared to the control-free case.
Maximization of the reachability set, on an appropriately chosen \Poincare section, allows for a greater space of potential transfer trajectories.
The use of the \Poincare section allows for design on a lower dimensional space and simplifies the design process.
Once an intersection is determined on the \Poincare section a transfer is calculated.

@@

@@

A periodic orbit is generated about the \( L_1 \) Lagrange point. 
A \Poincare section is chosen to allow for design on a lower dimensional subspace.
The section is defined along the \( \hat{x} \) axis, or defined with \( \alpha = \ang{0} \) and intersects both the initial and target orbits.
From the periodic orbit, a series of optimal trajectories are generated to approximate the reachable set.
The trajectories are generated from a fixed initial state of \( x_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix} \) over a fixed time span of \( t_f = 1.4 \).
By varying the angle \( \theta_d\) in~\cref{eq:constraints}, a different direction along the \Poincare section is maximized. 
The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:transfer_orbit}.
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{reach_trajectory} 
		\caption{Controlled Trajectories} \label{fig:reach_trajectory} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{manifold_trajectory} 
		\caption{Invariant Manifold} \label{fig:manifold_trajectory} 
	\end{subfigure}

	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{poincare_compare} 
		\caption{Poincar\`e Section} \label{fig:poincare_compare} 
	\end{subfigure} 
	\caption{Transfer Trajectory}
	\label{fig:transfer_orbit} 
\end{figure}
The optimal trajectories, under the influence of the control input \( \bar{u} \), are shown in~\cref{fig:reach_trajectory}.
Initially, the spacecraft is assumed to lie on the periodic orbit.
As a result, the intersection of this periodic orbit with the \Poincare section are two points corresponding to the two crossing of the orbit.
The use of the continuous low thrust propulsion expands the reachable set to region bounded by the red markers in~\cref{fig:poincare_compare}.
The reachable set is an ellipsoidal region with a major axis aligned along \( \theta \approx \ang{70} \).

@@

\subsection{Geostationary Transfer}
This example demonstrates the feasibility of iteratively computing the reachable set. 
In this manner, it is possible to design arbitrary transfers which require extended transfer durations.
Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional three body system. 
It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
The initial and final states of are illustrated in~\cref{fig:geo_transfer_target}.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{initial_final} % requires the graphicx package
   \caption{Transfer target}
   \label{fig:geo_transfer_target}
\end{figure}
In order to effectively utilize the dynamics of the three-body system the stable manifold of the periodic orbit is targeted.
This type of low energy transfer is best suited for autonomous spacecraft missions as the long time of flight is undesirable for manned spacecraft. 
Future proposals for permanent lunar spacecraft and bases will require frequent supply missions to remain viable. 
Low energy transfers from the Earth to the Lagrange points are necessary for future missions.

We generate the stable invariant manifold associated with the periodic orbit.
The stable manifold is propagated to the \Poincare section directed along the positive \( \hat{x} \) axis, or \( \alpha =\ang{0} \).
The reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section.
The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) defined on the \Poincare section.
This allows us to approximate the set of states that are achievable in the \( \hat{x} \text{ vs. } \hat{\dot{x}} \) space.
A final state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.
With each reachable set we move the final trajectory closer to the stable manifold.
\Cref{fig:geo_transfer_poincare} shows that after eight iterations the resulting trajectory intersects the stable manifold. 
Combining these trajectories results in the powered portion of the transfer from the geostationary orbit to the stable manifold. 
\Cref{fig:geo_transfer_full,fig:geo_transfer_zoom} shows the resulting trajectory, with the final trajectory shown in red which ensures the intersection with the stable invariant manifold.
Once at the stable manifold, no further control input is required and the vehicle will coast towards the target periodic orbit.
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_full} 
		\caption{Transfer} \label{fig:geo_transfer_full} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_zoom} 
		\caption{State Trajectory} \label{fig:geo_transfer_zoom} 
	\end{subfigure}

>>>>

%% Group authors per affiliation:
\author{Shankar Kulumani\fnref{kulumanifootnote}}
\address{George Washington University, 800 22nd St NW, Washington, DC 20052}

\author{Taeyoung Lee\fnref{leefootnote}}
\address{George Washington University, 800 22nd St NW, Washington, DC 20052}

@@

\begin{abstract}
A computational approach is developed for the design of continuous low thrust transfers in the planar circular restricted three-body problem.
The transfer design method of invariant manifolds is extended with the addition of continuous low thrust propulsion.
A reachable region is generated and it is used to determine transfer opportunities, analogous to the intersection of invariant manifolds.
The reachable set is developed on a lower dimensional \Poincare section and used to design transfer trajectories. 
%Computational geometric mechanics allows for the development of variational integrators and a discrete optimal control problem.
This is solved numerically as a discrete optimal control problem using a variational integrator.
This provides for a geometrically exact and numerically efficient method for the motion in the three-body problem.
Two numerical simulations are provided developing a transfer from a \( L_1 \) periodic orbit in  the Earth-Moon system to a target orbit about the Moon as well as a transfer from geostationary orbit to a \( L_1 \) periodic orbit.
\end{abstract}

@@

A discrete optimal control problem is formulated to determine the reachability set on a \Poincare section.
Given an initial condition and fixed time horizon, the reachable set is the set of states attainable, subject to the operational constraints of the spacecraft. 
Generation of this reachability set allows for the extension of the previous control-free missions in the three-body problem.
In addition, the generation of the reachable set allows for a more systematic method of determining initial conditions and eases the burden on the designer. 
Rather than relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on a \Poincare section.
This simple methodology allows for extended transfer trajectories which iteratively approach a target region.
The addition of low-thrust propulsion affords the ability to enlarge the reachable set as compared to the control-free case.
Maximization of the reachability set, on an appropriately chosen \Poincare section, allows for a greater space of potential transfer trajectories.
The use of the \Poincare section allows for design on a lower dimensional space and simplifies the design process.
Once an intersection is determined on the \Poincare section a transfer is calculated.

@@

\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{reach_trajectory} 
		\caption{Controlled Trajectories} \label{fig:reach_trajectory} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{manifold_trajectory} 
		\caption{Invariant Manifold} \label{fig:manifold_trajectory} 
	\end{subfigure}

	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{poincare_compare} 
		\caption{Poincar\`e Section} \label{fig:poincare_compare} 
	\end{subfigure} 
	\caption{Transfer Trajectory}
	\label{fig:transfer_orbit} 
\end{figure}

@@

A periodic orbit is generated about the \( L_1 \) Lagrange point. 
A \Poincare section is chosen to allow for design on a lower dimensional subspace.
The section is defined along the \( \hat{x} \) axis, or defined with \( \alpha = \ang{0} \) and intersects both the initial and target orbits.
From the periodic orbit, a series of optimal trajectories are generated to approximate the reachable set.
The trajectories are generated from a fixed initial state of \( x_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix} \) over a fixed time span of \( t_f = 1.4 \).
By varying the angle \( \theta_d\) in~\cref{eq:constraints}, a different direction along the \Poincare section is maximized. 
The intersection of the optimal trajectories as well as those of the target Moon orbit with the \Poincare section are shown in~\cref{fig:transfer_orbit}.

The optimal trajectories, under the influence of the control input \( \bar{u} \), are shown in~\cref{fig:reach_trajectory}.
Initially, the spacecraft is assumed to lie on the periodic orbit.
As a result, the intersection of this periodic orbit with the \Poincare section are two points corresponding to the two crossing of the orbit.
The use of the continuous low thrust propulsion expands the reachable set to region bounded by the red markers in~\cref{fig:poincare_compare}.
The reachable set is an ellipsoidal region with a major axis aligned along \( \theta \approx \ang{70} \).

@@

\subsection{Geostationary Transfer}
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{initial_final} % requires the graphicx package
   \caption{Transfer target}
   \label{fig:geo_transfer_target}
\end{figure}

This example demonstrates the feasibility of iteratively computing the reachable set. 
In this manner, it is possible to design arbitrary transfers which require extended transfer durations.
Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional three body system. 
It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
The initial and final states of are illustrated in~\cref{fig:geo_transfer_target}.

In order to effectively utilize the dynamics of the three-body system the stable manifold of the periodic orbit is targeted.
This type of low energy transfer is best suited for autonomous spacecraft missions as the long time of flight is undesirable for manned spacecraft. 
Future proposals for permanent lunar spacecraft and bases will require frequent supply missions to remain viable. 
Low energy transfers from the Earth to the Lagrange points are necessary for future missions.

We generate the stable invariant manifold associated with the periodic orbit.
The stable manifold is propagated to the \Poincare section directed along the positive \( \hat{x} \) axis, or \( \alpha =\ang{0} \) denoted by the horizontal black line.
The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section.
The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) in~\cref{eqn:constraints} defined on the \Poincare section.
This allows us to approximate the set of states that are achievable in the \( \hat{x} \text{ vs. } \hat{\dot{x}} \) space.
A final state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.
With each reachable set we move the final trajectory closer to the stable manifold.
\Cref{fig:geo_transfer_poincare} shows that after eight iterations the resulting trajectory intersects the stable manifold. 
Combining these trajectories results in the powered portion of the transfer from the geostationary orbit to the stable manifold. 
\Cref{fig:geo_transfer_full,fig:geo_transfer_zoom} shows the resulting trajectory, with the final trajectory shown in red which ensures the intersection with the stable invariant manifold.
Once at the stable manifold, no further control input is required and the vehicle will coast towards the target periodic orbit.
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_full} 
		\caption{Transfer} \label{fig:geo_transfer_full} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_zoom} 
		\caption{State Trajectory} \label{fig:geo_transfer_zoom} 
	\end{subfigure}
